---
title: "Getting started with Summary-statistic regional association scanning"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with Summary-statistic regional association scanning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
set.seed(35)
library(rasSS)
```

rasSS extends the RAS method to accept GWAS summary statistics. It replaces the per-window 
regressions that required individual-level data with 1-df weighted burden statistics, 
calculated by `t_burden = w'Z / sqrt(w'Rw)`, using the marginal Z-scores, weights
derived from a discovery set (marginal effect sizes), and an external LD reference. 
The resulting profile is then inputted into the original, unmodified RAS change-point 
detection. This vignette demonstrates a full example workflow on a small simulated chromosome.

## Simulate a chromosome and an LD reference

```{r simulate}
n_snps <- 200
R_true <- matrix(0, n_snps, n_snps)
for (b in 0:3) {
  idx <- b * 50 + 1:50
  R_true[idx, idx] <- outer(1:50, 1:50, function(i, j) 0.8^abs(i - j))
}

X_ref <- sim_genotypes(2000, R_true) # external reference panel
R_emp <- cor(X_ref)
mask  <- prune_ld(R_emp, thresh = 0.2) # LD pruning mask
```

## Discovery and target summary statistics

```{r stats}
beta <- rep(0, n_snps)
beta[101:110] <- 0.5 # one sparse causal cluster

X_d <- sim_genotypes(1000, R_true)
X_t <- sim_genotypes(1000, R_true)
y_d <- make_pheno(X_d, beta, "continuous", seed = 2)
y_t <- make_pheno(X_t, beta, "continuous", seed = 3)

disc <- get_marginal_stats(X_d, y_d) # discovery weights w
targ <- get_marginal_stats(X_t, y_t) # target Z-scores Z
```

## Scan with explicit window parameters

```{r scan}
scan <- scan_ss(disc$beta, targ$z, R_emp, mask,
                skip1 = 5, skip2 = 5,
                min_window_size = 5, max_window_size = 15)

plot(scan$x, scan$y, type = "l",
     xlab = "SNP position", ylab = expression(-log[10](p)),
     main = "rasSS summary-statistic profile")
abline(v = 105, lty = 2, col = "red")  # true causal cluster
```

## The burden statistic for a single window

```{r burden}
win <- 96:115
t_burden(w = disc$beta[win], Z = targ$z[win], R = R_emp[win, win])
```

## Peak detection with the original RAS CPD

rasSS does not modify the original detection algorithm; the profile generated above 
is inputted into RAS's own `ras_detect()` / `ras_validate()`, specifying all detection parameters.

```{r detect}
cp <- quiet(RAS::ras_detect(x = seq_along(scan$x), y = scan$y,
                            window_size = 12, slope_check_window_size = 8,
                            slope.p.values.threshold.left = 1e-4,
                            slope.p.values.threshold.right = 1e-4))
val <- quiet(RAS::ras_validate(this.result = cp, x = seq_along(scan$x), y = scan$y,
                               this.start = 1, this.skip = 5,
                               second_window_size = 20, min_signal = 2.5,
                               p.value.threshold = 1e-3))
scan$x[val$tau_hats] # detected peak positions (SNP coordinates)
```

## Binary traits

For case-control data, replace the marginal statistics with the Rao score version;
everything else is identical as follows.

```{r binary}
y_b   <- make_pheno(X_t, beta, "binary", seed = 4)
stat_b <- get_marginal_stats_bin(X_t, y_b)
head(stat_b$z, 3)
```

The validation studies (paired comparison against the individual-level method,
type I calibration, and the paper replication) can be found on the pkgdown website
under Articles.